---
title: "Simulation"
author: "Anonymous"
date: "2026-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(patchwork)
library(ggdag)
```

```{r configuration}
granularity <- 101
```

## Setup

There is no native way of calculating the cumulative distribution function (CDF) in R.
Hence, we define our own custom function.

```{r calc-cdf}
cdf.beta <- function(alpha = 2, beta = 2, x = NA) {
  # calculate the probability density function
  pdf <- dbeta(seq(from = 0, to = 1, length.out = granularity), alpha, beta)

  # calculate the cumulative distribution function
  cdf <- c(0)
  for (i in 2:101) {
    cdf[i] <- cdf[i-1] + pdf[i]
  }
  cdf <- cdf/100
  
  if(is.na(x)) {
    return(cdf)
  } else {
    return(cdf[(granularity-1)*x + 1])
  }
}
```

The following helper function allows plotting the CDF of a Beta-distribution.

```{r visualize-cdf}
plot.cdf <- function(cdf) {
  plt <- data.frame(
    x = seq(0, 1, length.out = granularity),
    y = cdf
  ) %>% 
    ggplot(aes(x = x, y = y)) +
      geom_line() +
      labs(x = "Resources", y = "Information density") +
      ylim(0, 1) +
      theme_bw()
  
  return(plt)
}
```

```{r example}
plot.cdf(cdf.beta(alpha = 5, beta = 2))
```

## Application

The following subsections illustrate the application of the simulation framework.

### Chain

```{r chain-dag}
dag <- dagify(
  reng ~ customer,
  srs ~ reng,
  
  exposure = "customer", outcome = "srs",
  labels = c(customer = "Customer", reng = "Requirements Engineer", 
             srs = "Requirements Specification"),
  
  coords = list(
    x = c(customer = 0, reng = 1, srs = 2),
    y = c(customer = 0, reng = 0, srs = 0)
  )
)

# visualize the DAG
ggdag_status(dag, use_labels = "label", text = FALSE) +
  guides(fill = "none", color = "none") + 
  theme_dag()
```

```{r chain-components}
plt.reng <- plot.cdf(cdf.beta(alpha = 2, beta = 5))
plt.srs <- plot.cdf(cdf.beta(alpha = 5, beta = 2))

plt.reng | plt.srs
```

### Joint node

The following case illustrates a **joint**, i.e., a node where two *alternative* paths collide and the node needs to decide, which path to take.
Naturally, the node will choose the path based on efficiency, i.e., the relationship between information output and cost.
In the example, we consider two paths:

1. Formal path: Customer $\rightarrow$ Requirements Engineer $\rightarrow$ Requirements Specification $\rightarrow$ Developer
2. Informal path: Customer $\rightarrow$ Developer

```{r fork-dag}
dag <- dagify(
  reng ~ customer,
  srs ~ reng,
  dev ~ srs + customer,
  
  exposure = "customer", outcome = "dev",
  labels = c(customer = "Customer", reng = "Requirements Engineer", 
             srs = "Requirements Specification", dev = "Developer"),
  
  coords = list(
    x = c(customer = 0, reng = 1, srs = 2, dev = 3),
    y = c(customer = 0, reng = 1, srs = 1, dev = 0.5)
  )
)

# visualize the DAG
ggdag_status(dag, use_labels = "label", text = FALSE) +
  guides(fill = "none", color = "none") + 
  theme_dag()
```

Each transfer of information $x \rightarrow y$ is defined by a transfer function:

$$f(i: \text{information}, c: \text{complexity}, s: \text{skill}, t: \text{time}) \rightarrow i: \text{information}$$

For simplicity, we reduce the number of influencing factors to four:

1. **Input information** $i \in [0; 1]$, the information provided by the sender
2. **Task complexity** $c \in \mathbb{R}^+$, the baseline challenge of the performed activity
3. **Stakeholder skill** $s \in \mathbb{R}^+$, the ability of the stakeholder performing the activity
4. **Time spent** $t \in [0; 1]$, the amount of resources (here measured in hours) invested into the activity

The function $f(i, c, s, t)$ is then realized as the cumulative distribution function of a beta-distribution:

$$f(i, c, s, t) = \text{CDF}_\beta(\alpha = c, \beta = s, t) * i$$

This way, higher values of complexity $c$ will make transferring more difficult, but higher values of skill $s$ make it easier.
The more time $t$ spent on the activity, the more information will be transferred.
Finally, the input information $i$ defines the upper bound of information to transfer.
The following figure illustrates the relation between the variables.

```{r functional-definition}
f2 <- function(i, c, s, t) {
  return(cdf.beta(alpha = c, beta = s, x = t)*i)
}

c.high <- 6
c.low <- 2
s.low <- 1
s.high <- 5

p1 <- plot.cdf(cdf.beta(alpha = c.low, beta = s.low)) + ggtitle("Low skill, low complexity")
p2 <- plot.cdf(cdf.beta(alpha = c.low, beta = s.high)) + ggtitle("High skill, low complexity")
p3 <- plot.cdf(cdf.beta(alpha = c.high, beta = s.low)) + ggtitle("Low skill, high complexity")
p4 <- plot.cdf(cdf.beta(alpha = c.high, beta = s.high)) + ggtitle("High skill, high complexity")

(p1 | p2) / (p3 | p4)
```

Now we can simulate the case above in two scenarios: waterfall and agile.
The main difference is that in the waterfall scenario, the complexity of the *direct communication* (customer $\rightarrow$ developer) is much higher than in the agile scenario.

```{r scenario-waterfall}
# configuration
info.customer <- 0.8
# elicitation activity (customer -> req engineer)
c.elicitation <- 6
s.elicitation <- 5
t.elicitation <- 0.8
# specification activity (req engineer -> specification)
c.specification <- 4
s.specification <- 7
t.specification <- 0.7
# reading activity (specification -> developer)
c.reading <- 2
s.reading <- 4
t.reading <- 0.6
# direct communication activity (customer -> developer)
c.direct <- 12
s.direct <- 3
t.direct <- 0.9

# simulation
info.elicitation <- f2(i = info.customer, c = c.elicitation, s = s.elicitation, t = t.elicitation)
info.specification <- f2(i = info.elicitation, c = c.specification, s = s.specification, t = t.specification)
info.reading <- f2(i = info.specification, c = c.reading, s = s.reading, t = t.reading)
info.reading

info.communicating <- f2(i = info.customer, c = c.direct, s = s.direct, t = t.direct)
info.communicating

# visualization
plt.customer2reng <- plot.cdf(cdf.beta(alpha = c.elicitation, beta = s.elicitation) * info.customer) + 
  geom_vline(xintercept = t.elicitation, color = "red", linetype = "dashed") + 
  geom_hline(yintercept = info.customer, color = "grey", linetype = "dashed") + 
  ggtitle("Customer -> Req Engineer")
plt.reng2srs <- plot.cdf(cdf.beta(alpha = c.specification, beta = s.specification) * info.specification) + 
  geom_vline(xintercept = t.specification, color = "red", linetype = "dashed") +  
  ggtitle("Req Engineer -> Req Specification")
plt.srs2dev <- plot.cdf(cdf.beta(alpha = c.reading, beta = s.reading) * info.reading) + 
  geom_vline(xintercept = info.reading, color = "red", linetype = "dashed") + 
  ggtitle("Req Specification -> Developer")

plt.customer2dev <- plot.cdf(cdf.beta(alpha = c.direct, beta = s.direct) * info.customer) +
  geom_vline(xintercept = t.direct, color = "red", linetype = "dashed") + 
  ggtitle("Customer -> Developer")

(plt.customer2reng | plt.reng2srs) / (plt.srs2dev | plt.customer2dev)
```

```{r scenario-waterfall}
# configuration
info.customer <- 0.8
# elicitation activity (customer -> req engineer)
c.elicitation <- 6
s.elicitation <- 5
t.elicitation <- 0.8
# specification activity (req engineer -> specification)
c.specification <- 4
s.specification <- 7
t.specification <- 0.7
# reading activity (specification -> developer)
c.reading <- 2
s.reading <- 4
t.reading <- 0.7
# direct communication activity (customer -> developer)
c.direct <- 5
s.direct <- 3
t.direct <- 0.9

# simulation
info.elicitation <- f2(i = info.customer, c = c.elicitation, s = s.elicitation, t = t.elicitation)
info.specification <- f2(i = info.elicitation, c = c.specification, s = s.specification, t = t.specification)
info.reading <- f2(i = info.specification, c = c.reading, s = s.reading, t = t.reading)
info.reading
info.communicating <- f2(i = info.customer, c = c.direct, s = s.direct, t = t.direct)
info.communicating

# visualization
plt.customer2reng <- plot.cdf(cdf.beta(alpha = c.elicitation, beta = s.elicitation) * info.customer) + 
  geom_vline(xintercept = t.elicitation, color = "red", linetype = "dashed") +
  geom_hline(yintercept = info.customer, color = "grey", linetype = "dashed") +  
  ggtitle("Customer -> Req Engineer")
plt.reng2srs <- plot.cdf(cdf.beta(alpha = c.specification, beta = s.specification) * info.specification) + 
  geom_vline(xintercept = t.specification, color = "red", linetype = "dashed") +  
  ggtitle("Req Engineer -> Req Specification")
plt.srs2dev <- plot.cdf(cdf.beta(alpha = c.reading, beta = s.reading) * info.reading) + 
  geom_vline(xintercept = t.reading, color = "red", linetype = "dashed") + 
  ggtitle("Req Specification -> Developer")

plt.customer2dev <- plot.cdf(cdf.beta(alpha = c.direct, beta = s.direct) * info.communicating) +
  geom_vline(xintercept = t.direct, color = "red", linetype = "dashed") + 
  ggtitle("Customer -> Developer")

(plt.customer2reng | plt.reng2srs) / (plt.srs2dev | plt.customer2dev)
```

### old

```{r functional-definition}
f <- function(i, c, s, t) {
  return(cdf.beta(max(0, c - s), t, i))
}

c <- 6 # baseline complexity
s.low <- 1
s.high <- 5

t.low <- 2
t.high <- 5

p1 <- plot.cdf(cdf.beta(alpha = c - s.low, beta = t.low)) + ggtitle("Little time, low skill")
p2 <- plot.cdf(cdf.beta(alpha = c - s.high, beta = t.low)) + ggtitle("Little time, high skill")
p3 <- plot.cdf(cdf.beta(alpha = c - s.low, beta = t.high)) + ggtitle("Much time, low skill")
p4 <- plot.cdf(cdf.beta(alpha = c - s.high, beta = t.high)) + ggtitle("Much time, high skill")

(p1 | p2) / (p3 | p4)
```

```{r scenario-waterfall}
# configuration
info.customer <- 0.8
# elicitation activity (customer -> req engineer)
c.elicitation <- 6
s.elicitation <- 3
t.elicitation <- 2
# specification activity (req engineer -> specification)
c.specification <- 4
s.specification <- 2
t.specification <- 2
# reading activity (specification -> developer)
c.reading <- 2
s.reading <- 1
t.reading <- 3
# direct communication activity (customer -> developer)
c.direct <- 12
s.direct <- 1
t.direct <- 4

# simulation
info.elicitation <- f(i = info.customer, c = c.elicitation, s = s.elicitation, t = t.elicitation)
info.specification <- f(i = info.elicitation, c = c.specification, s = s.specification, t = t.specification)
info.reading <- f(i = info.specification, c = c.reading, s = s.reading, t = t.reading)
info.reading
info.communicating <- f(i = info.customer, c = c.direct, s = s.direct, t = t.direct)
info.communicating

# visualization
plt.customer2reng <- plot.cdf(cdf.beta(alpha = c.elicitation-s.elicitation, beta = t.elicitation)) + 
  geom_vline(xintercept = info.customer, color = "red", linetype = "dashed") + 
  ggtitle("Customer -> Req Engineer")
plt.reng2srs <- plot.cdf(cdf.beta(alpha = c.specification-s.specification, beta = t.specification)) + 
  geom_vline(xintercept = info.elicitation, color = "red", linetype = "dashed") +  
  ggtitle("Req Engineer -> Req Specification")
plt.srs2dev <- plot.cdf(cdf.beta(alpha = c.reading-s.reading, beta = t.reading)) + 
  geom_vline(xintercept = info.specification, color = "red", linetype = "dashed") + 
  ggtitle("Req Specification -> Developer")

plt.customer2dev <- plot.cdf(cdf.beta(alpha = c.direct-s.direct, beta = t.direct)) +
  geom_vline(xintercept = info.customer, color = "red", linetype = "dashed") + 
  ggtitle("Customer -> Developer")

(plt.customer2reng | plt.reng2srs) / (plt.srs2dev | plt.customer2dev)
```

```{r scenario-agile}
# configuration
info.customer <- 0.8
# elicitation activity (customer -> req engineer)
c.elicitation <- 6
s.elicitation <- 3
t.elicitation <- 2
# specification activity (req engineer -> specification)
c.specification <- 4
s.specification <- 2
t.specification <- 2
# reading activity (specification -> developer)
c.reading <- 2
s.reading <- 1
t.reading <- 3
# direct communication activity (customer -> developer)
c.direct <- 5
s.direct <- 1
t.direct <- 2

# simulation
info.elicitation <- f(i = info.customer, c = c.elicitation, s = s.elicitation, t = t.elicitation)
info.specification <- f(i = info.elicitation, c = c.specification, s = s.specification, t = t.specification)
info.reading <- f(i = info.specification, c = c.reading, s = s.reading, t = t.reading)
info.reading
info.communicating <- f(i = info.customer, c = c.direct, s = s.direct, t = t.direct)
info.communicating

# visualization
plt.customer2reng <- plot.cdf(cdf.beta(alpha = c.elicitation-s.elicitation, beta = t.elicitation)) + 
  geom_vline(xintercept = info.customer, color = "red", linetype = "dashed") + 
  ggtitle("Customer -> Req Engineer")
plt.reng2srs <- plot.cdf(cdf.beta(alpha = c.specification-s.specification, beta = t.specification)) + 
  geom_vline(xintercept = info.elicitation, color = "red", linetype = "dashed") +  
  ggtitle("Req Engineer -> Req Specification")
plt.srs2dev <- plot.cdf(cdf.beta(alpha = c.reading-s.reading, beta = t.reading)) + 
  geom_vline(xintercept = info.specification, color = "red", linetype = "dashed") + 
  ggtitle("Req Specification -> Developer")

plt.customer2dev <- plot.cdf(cdf.beta(alpha = c.direct-s.direct, beta = t.direct)) +
  geom_vline(xintercept = info.customer, color = "red", linetype = "dashed") + 
  ggtitle("Customer -> Developer")

(plt.customer2reng | plt.reng2srs) / (plt.srs2dev | plt.customer2dev)
```
